Pyg4ometry: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-09-01 public/paper/pyg4ometry.tex
more fluka2gdml paper bits

\section{FLUKA to GDML conversion}
FLUKA geometry is based upon a limited set of primitives (referred to
as \textit{bodies} which can be combined using Boolean operations.  A
\textit{zone} consists of one or more bodies or \textit{subzones}
combined using intersections and subtractions.  Zones may then be
further combined using union operations to form \textit{regions},
which is defined as the union of one or more zones, as well as a
material.

@@

The FLUKA CSG ASCII format is parsed using an ANTLR4-generated parser,
producing an abstract syntax tree (AST).  The resulting AST is then walked
to populate Region instances with zones and bodies.  With the region
instances populated they can then be manipulated and translated into GDML.
The translation involves a number of special steps to bridge the two
disparate formats and ensure the resulting GDML is well-formed ans usable
in Geant4.  Some of these steps are simple, for example in FLUKA unions can
be disconnected, but in Geant4 specifically only MultiUnions can be
disconnected, such that MultiUnions are used throughout the converted
geometry instead of the usual binary unions.  Other procedures are more
involved and are discussed in the rest of this section.

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/faradayCup2.pdf}
\caption{Example conversion of a simple FLUKA geometry to GDML, left panel :  
the original FLUKA geometry displayed in FLAIR, right panel : the GDML geometry 
viewed in pyg4ometry. The example is a Faraday cup used to capture 
and measure accelerator beam charge.}
\label{fig:region-ast}
\end{center}
\end{figure*}

@@

\subsection{Discussion}
A disjoint union in FLUKA is a union of two or more zones that are not
connected.  This is a valid construct in FLUKA but is explicitly
forbidden in Geant4 and must be removed during the conversion
process.  Disjoint unions are detected by checking for overlaps
between all zones within a union.

\begin{enumerate}
\item disjoint subtractions?
\item quadric specialisation
\item second order lattice
\item Materials 
\end{enumerate}

>>>>

\section{FLUKA to GDML conversion}
FLUKA geometry is based upon a limited set of primitives (referred to as
\textit{bodies}) which can be combined using Boolean operations. A
\textit{zone} consists of one or more bodies or \textit{subzones} combined
using intersections and subtractions.  Zones may then be further combined
using union operations to form \textit{regions}, which is defined as the
union of one or more zones, as well as a material.

@@

The FLUKA CSG ASCII format is parsed using an ANTLR4-generated parser,
producing an abstract syntax tree (AST).  The resulting AST is then walked
to populate Region instances with zones and bodies.  With the region
instances populated they can then be manipulated and translated into GDML.
The translation involves a number of special steps to bridge the two
disparate formats and ensure the resulting GDML is well-formed and usable
in Geant4.  Some of these steps are simple, for example in FLUKA unions can
be disconnected, but in Geant4 specifically only MultiUnions can be
disconnected, such that MultiUnions are used throughout the converted
geometry instead of the usual binary unions.  Other procedures are more
involved and are discussed in the rest of this section.

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/faradayCup2.pdf}
\caption{Example conversion of a simple FLUKA geometry to GDML, left panel :
the original FLUKA geometry displayed in FLAIR, right panel : the GDML geometry
viewed in pyg4ometry. The example is a Faraday cup used to capture
and measure accelerator beam charge.}
\label{fig:fluka-to-geant4-cup}
\end{center}
\end{figure*}

@@

\subsection{Discussion}

Figure~\ref{fig:fluka-to-geant4-cup} shows a Faraday cup implemented in
FLUKA and translated to GDML using pyg4ometry.  The conversion is clearly
accurate well and also includes the translation of the materials to GDML.
Many of the steps described above were directly applied in this model and
all the features are tested and demonstrated in the repository.  This set
of algorithms for bridging FLUKA with GDML covers a very broad range of
geometries, however a number of possible improvements for the future
persist.  For example, Boolean solids in FLUKA can in general be
disconnected, and this most typically manifests itself in the form of
disconnected unions, but is also allowed for intersections and
subtractions.  Disconnected unions are readily availabe in GDML with the
use of the MultiUnion solid type, however there is no valid means to
construct a disconnected subtraction or intersection in GDML.  One possible
solution would be to detect these two cases, and if found, split them into
their constituent parts and place separately as TesselatedSolid instances.

Furthermore, the quadric surface conversion can be further improved by
specialising on some of the individual forms of the quadric.  Simply
converting every quadric into a TesselatedSolid comes at a potential
performance cost in the tracking, as well as usability in pyg4ometry as the
user must provide an AABB.  In some cases tesselation in unavoidable (for
example a hyperbolic parabaloid), but parabolic cylinders could for example
be translated to an ExtrudedSolid.  This could provide both a performance
improvement in the tracking time and make pyg4ometry easier to use as an
AABB would not need to be provided beforehand.  This has not been
implemented as quadrics are relatively rarely used, but where quadrics are
used its often parabolic cylinders (e.g. magnet pole tips) this
specicalisation in particular would be worth implementing.
